# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
Task 1: mean and median total steps per day
------

#### Get averages by day
```{r}
setwd("/Users/testo/datasci/reprod/ass2/")
steps<-read.csv("activity.csv")
stepdays<-tapply(steps$steps, steps$date, sum)
hist(stepdays, main="Steps per day", xla="steps", yla="number of days")
```


## What is mean total number of steps taken per day?

#### Mean and median: 
```{r}
mean(stepdays, na.rm=T)
median(stepdays, na.rm=T)
```


## What is the average daily activity pattern?

### Plot time series:
```{r}
stepmin<-aggregate(steps$steps, list(steps$interval), mean, na.rm=T)
names(stepmin)<-c("interval", "steps")
plot(stepmin$interval, stepmin$steps, type="l")
```

### Get the interval with the most steps:
```{r}
stepmin[order(stepmin$steps),][288,]$interval
```

## Imputing missing values

### number of missing data points:

```{r}
length(which(is.na(steps$steps)))
```

### steps ~ time model:

```{r}
#note: R incorrectly thinks the HHMM time is an integer, but we can work with that instead of having to convert.
steps$hour = floor(steps$interval/100)
steps$period=floor(steps$hour/3)
steps$period<-factor(steps$period)
levels(steps$period)<-c("0-2", "3-5", "6-8", "9-11", "12-14", "15-17", "18-20", "21-23")
mod<-lm(steps ~ period, data=steps)
mod
```

Now imputing:

```{r}
steps$stepsi<-steps$steps
steps$stepsi[is.na(steps$steps)]<-predict(mod, newdata=steps[is.na(steps$steps),])
```

```{r}
stepdaysi<-tapply(steps$stepsi, steps$date, sum, na.rm=T)
stepdaysi
hist(stepdaysi, main="Steps per day (with imputed data)", xla="steps", yla="number of days", col="#ff99ff")
```
*mean (with imputed data)*:
```{r}
mean(stepdaysi, na.rm=T)
```
*median  (with imputed data)*:
```{r}
median(stepdaysi, na.rm=T)
```



## Are there differences in activity patterns between weekdays and weekends?


```{r}
steps$ddate<-as.character(steps$date)
steps$ddate<-as.Date(steps$ddate, format="%Y-%m-%d")
steps$weekday<-weekdays(steps$ddate)
steps$weekend<-F
steps$weekend[steps$weekday %in% c("Saturday", "Sunday")]<-T

stepmin.i.weekdays<-aggregate(steps$stepsi[!steps$weekend], list(steps$interval[!steps$weekend]), mean, na.rm=T)
stepmin.i.weekends<-aggregate(steps$stepsi[steps$weekend], list(steps$interval[steps$weekend]), mean, na.rm=T)
names(stepmin.i.weekdays)<-c("interval", "steps")
names(stepmin.i.weekends)<-c("interval", "steps")


par(mfrow = c(2,1), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0))
plot(stepmin.i.weekends$interval, stepmin.i.weekends$steps, pch="", ylab="Steps", xlab="", main="weekend", type="l", ylim=c(0,220), col="blue")
plot(stepmin.i.weekdays$interval, stepmin.i.weekdays$steps, pch="", ylab="Steps", xlab="", main="weekday", type="l",  ylim=c(0,220), col="darkred")
```
